### 2019-01-25 report/paper.tex
Add report files.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{COMP 6721: Project Report}
\author{Martin Khannouz}
%
%\authorrunning{COMP 472 [or COMP 6721] F. Author et al.}
\institute{40093517 \email{martin.khannouz@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref1}, an LNCS chapter~\cite{ref2}, a
book~\cite{ref1}, proceedings without editors~\cite{ref1},
and a homepage~\cite{ref_2}. Multiple citations are grouped
\cite{ref1,ref2}.
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{paper}
\end{document}

% vim: tw=50 ts=2

### 2019-01-25 src/engine.py
An engine that can print itself and execute moves. And check if a move is possible.

@@ -1,7 +1,8 @@
+import sys
 import numpy as np
 
 class Engine:
-    def __init__(self, width, height=8, card_count=24, max_turn=60):
+    def __init__(self, width=8, height=12, card_count=24, max_turn=60):
         self.width = width
         self.height = height
         self.board = np.zeros((width, height))
@@ -19,22 +20,22 @@ class Engine:
 
 	# Create list of position to check before adding a new card
         if move.type in {1, 3, 5, 7}: # horizontal move
-            placement_pos = np.array([move.pos, [move.pos[0]+1, move.pos[1]]])
+            placement_pos = np.array([move.pos, (move.pos[0]+1, move.pos[1])])
         if move.type in {2, 4, 6, 8}: # vertical move
-            placement_pos = np.array([move.pos, [move.pos[0], move.pos[1]+1]])
+            placement_pos = np.array([move.pos, (move.pos[0], move.pos[1]+1)])
         if move.type in {1, 3, 5, 7}: # horizontal move
-            no_empty_pos = np.array([[move.pos[0], move.pos[1]-1], [move.pos[0]+1, move.pos[1]-1]])
+            no_empty_pos = np.array([(move.pos[0], move.pos[1]-1), (move.pos[0]+1, move.pos[1]-1)])
         if move.type in {2, 4, 6, 8}: # vertical move
-            no_empty_pos = np.array([[move.pos[0], move.pos[1]-1]])
+            no_empty_pos = np.array([(move.pos[0], move.pos[1]-1)])
         if move.recycling:
             if not self.is_on_board(move.pos_rec) or self.cards[move.pos_rec] == 0: #Card to move is outside of the map or there is no card
                 return False
             card_to_rec = self.cards[move.pos_rec]
             if card_to_rec in {1, 3, 5, 7}: # Horizontal recycling
                 #The two cells below the move *and* the two cells above the recycling card have to be *not* be empty
-                empty_pos = np.array([[move.pos_rec[0], move.pos_rec[1]+1], [move.pos_rec[0]+1, move.pos_rec[1]+1]])
+                empty_pos = np.array([(move.pos_rec[0], move.pos_rec[1]+1), (move.pos_rec[0]+1, move.pos_rec[1]+1)])
             elif card_to_rec in {2, 4, 6, 8}: # vertical recycling and vertical move
-                empty_pos = np.array([[move.pos_rec[0], move.pos_rec[1]+2]])
[…]

### 2019-01-25 src/main.py

@@ -1,0 +1,19 @@
+from engine import Engine
+from move import Move
+
+engine = Engine()
+move = Move()
+move.recycling = False
+move.type = 1
+move.pos = (0,0)
+val = engine.execute(move)
+move.type = 6
+move.pos = (2,0)
+val = engine.execute(move)
+move.type = 6
+move.pos = (2,2)
+val = engine.execute(move)
+print(val)
+print(engine.board)
+engine.print()
+print(engine.board)

### 2019-01-25 src/move.py

@@ -1,0 +1,6 @@
+class Move:
+    def __init__(self):
+        self.recycling = False
+        self.type = 0
+        self.pos = (-1,-1)
+        self.pos_rec = (-1,-1)

### 2019-01-25 tests/test_engine.py
Add test file.

@@ -1,0 +1,43 @@
+from engine import Engine
+from move import Move
+
+def test_move_wrong_recycling():
+    engine = Engine()
+    move = Move()
+    move.recycling = True
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == False)
+
+def test_move_right():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+
+def test_move_duplicate():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    val = engine.execute(move)
+    assert(val == False)
+
+def test_move_overlap():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    move.pos = (1,0)
+    val = engine.execute(move)
+    assert(val == False)

### 2019-01-25 src/engine.py
Add thing in engine.py.

@@ -1,5 +1,5 @@
-import sys
 import numpy as np
+from move import Move
 
 class Engine:
     def __init__(self, width=8, height=12, card_count=24, max_turn=60):
@@ -93,6 +93,7 @@ class Engine:
             self.card_count -= 1
         # TODO keep a list of move done so far
         return True
+
     def is_on_board(self, pos):
         if pos[0] < 0 or pos[0] > self.width:
             return False
@@ -126,18 +127,29 @@ class Engine:
                     line = line + bg_white + fg_black + "x" + bg_default + fg_default
                 elif self.board[pos] == 4:
                     line = line + bg_white + fg_black + "o" + bg_default + fg_default
-            sys.stdout.write(line + "\n")
-            # print(line)
+            print(line)
 
+    def isWinning(self):
+        '''Return who is winning.
+        -1: nobody
+         0: player 1
+         1: player 2
+         2: tie
+        '''
+        return -1
     def play(self, player1, player2):
         self.ais = [player1, player2]
 
         current_player = 0
         current_turn = 0
         while True:
-            move = self.ais[current_player].play(self)
+            move = self.ais[current_player].play(self) #Call the player function
             print("Player ", current_player, ": ", move)
             legal_move = self.execute(move)
             if not legal_move:
                 continue
             current_player = (current_player+1)%2
+
+            who_win = self.isWinning()
+            if who_win >= 0:
+                return who_win

### 2019-01-25 src/main.py

@@ -17,3 +17,4 @@ print(val)
 print(engine.board)
 engine.print()
 print(engine.board)
+# export PYTHONPATH=$PYTHONPATH:`pwd`/src

### 2019-01-25 src/engine.py
Add tests and fix the engine accordingly.

@@ -31,33 +31,57 @@ class Engine:
             if not self.is_on_board(move.pos_rec) or self.cards[move.pos_rec] == 0: #Card to move is outside of the map or there is no card
                 return False
             card_to_rec = self.cards[move.pos_rec]
+            if move.pos_rec == move.pos and card_to_rec == move.type:
+                return False
             if card_to_rec in {1, 3, 5, 7}: # Horizontal recycling
-                #The two cells below the move *and* the two cells above the recycling card have to be *not* be empty
+                #The two cells above the recycling card have to be empty
                 empty_pos = np.array([(move.pos_rec[0], move.pos_rec[1]+1), (move.pos_rec[0]+1, move.pos_rec[1]+1)])
-            elif card_to_rec in {2, 4, 6, 8}: # vertical recycling and vertical move
+                recycling_pos = np.array([move.pos_rec, (move.pos_rec[0]+1, move.pos_rec[1])])
+            elif card_to_rec in {2, 4, 6, 8}: # vertical recycling
                 empty_pos = np.array([(move.pos_rec[0], move.pos_rec[1]+2)])
+                recycling_pos = np.array([move.pos_rec, (move.pos_rec[0], move.pos_rec[1]+1)])
         else:
             empty_pos = np.array([])
+            recycling_pos = np.array([])
 
         # From now on, the empty_pos array contains all position that should be either empty or outside the board (to recycle).
         # no_empty_pos contains all positions that should be either filled or outside the board (to place above)
         # placement_pos is where the card will be placed, thus, all position in this array should be empty and on the board.
+        # recycling_pos contains the current position of the card to recycle
 
         for i in range(placement_pos.shape[0]):
             pos = tuple(placement_pos[i])
-            if not self.is_on_board(pos) or self.board[pos] != 0:
+            #If we recycle a card partially on itself, we need to consider the recycled cells as empty
+            if not self.is_on_board(pos) or (self.board[pos] != 0 and pos not in recycling_pos):
                 return False
         for i in range(no_empty_pos.shape[0]):
             pos = tuple(no_empty_pos[i])
-            if self.is_on_board(pos) and self.board[pos] == 0: #on board and empty
+            on_board = self.is_on_board(pos)
+            is_empty = self.board[pos] == 0
+            #If we recycle a card to put it above, we need to consider the cells recycled as empty
[…]

### 2019-01-25 tests/test_engine.py

@@ -10,6 +10,93 @@ def test_move_wrong_recycling():
     val = engine.execute(move)
     assert(val == False)
 
+def test_recycling():
+    '''Test replacing somewhere else'''
+    engine = Engine()
+    engine.card_count = 1
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    move.recycling = True
+    move.type = 6
+    move.pos = (2,0)
+    move.pos_rec = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+
+def test_recycling_same_cells1():
+    '''Test recycling on the same cells used before'''
+    engine = Engine()
+    engine.card_count = 1
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    engine.print()
+    move.recycling = True
+    move.type = 6
+    move.pos = (1,0)
+    move.pos_rec = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+
+def test_recycling_same_cells2():
+    '''Test recycling on the same cells used before with same orientation'''
+    engine = Engine()
+    engine.card_count = 1
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    engine.print()
+    move.recycling = True
+    move.type = 7
+    move.pos = (0,0)
+    move.pos_rec = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+
+def test_recycling_same_cells3():
+    '''Test recycling with the same move'''
+    engine = Engine()
+    engine.card_count = 1
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    move.recycling = True
+    move.type = 1
+    move.pos = (0,0)
+    move.pos_rec = (0,0)
+    val = engine.execute(move)
+    assert(val == False)
+
+def test_recycling_above():
+    '''Test recycling above the current'''
+    engine = Engine()
+    engine.card_count = 1
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    val = engine.execute(move)
+    assert(val == True)
+    engine.print()
+    move.recycling = True
[…]

### 2019-01-26 src/engine.py
Add an all_lines function.

@@ -9,8 +9,128 @@ class Engine:
         self.cards = np.zeros((width, height))
         self.card_count = card_count
         self.max_turn = max_turn
+        self.previous_moves = []
+        self.win_length = 4
+        #matching_cells describes how cells can form lines together.
+        #for instance, a cell with 1 (red with filled dot) share the same color as 2 (red with empty dot)
+        #matching_cells[0] is the matrix for the color
+        self.matching_cells = np.zeros((2, 4, 4))
+        self.matching_cells[0][0][0] = 1
+        self.matching_cells[0][1][1] = 1
+        self.matching_cells[0][2][2] = 1
+        self.matching_cells[0][3][3] = 1
+        self.matching_cells[0][0][1] = 1
+        self.matching_cells[0][1][0] = 1
+        self.matching_cells[0][2][3] = 1
+        self.matching_cells[0][3][2] = 1
+        #matching_cells[1] is the matrix for the dot
+        self.matching_cells[1][0][0] = 1
+        self.matching_cells[1][1][1] = 1
+        self.matching_cells[1][2][2] = 1
+        self.matching_cells[1][3][3] = 1
+        self.matching_cells[1][0][2] = 1
+        self.matching_cells[1][2][0] = 1
+        self.matching_cells[1][1][3] = 1
+        self.matching_cells[1][3][1] = 1
+
+        self.all_lines = self.build_all_lines()
+
+    def build_all_lines(self):
+        '''This function build all lines of the board with coordinates of each cell. Row, column, and diagonals'''
+        board = self.board
+        shape = board.shape
+        all_lines = []
+        for x in range(shape[0]):
+            all_lines.append([])
+            for y in range(shape[1]):
+                all_lines[-1].append((x,y))
+        for y in range(shape[1]):
+            all_lines.append([])
+            for x in range(shape[0]):
+                all_lines[-1].append((x,y))
+
+        for y in range(shape[1]):
+            x = 0
+            l = self.get_counter_diagonal_line(x,y)
+            if len(l) >= self.win_length:
+                all_lines.append(l)
+            l = self.get_diagonal_line(x,y)
+            if len(l) >= self.win_length:
+                all_lines.append(l)
+
+        for x in range(1,shape[0]):
+            l = self.get_counter_diagonal_line(x,0)
+            if len(l) >= self.win_length:
+                all_lines.append(l)
+            l = self.get_diagonal_line(x,shape[1]-1)
[…]

### 2019-01-26 src/engine.py
Add test for winning conditions. Split tests for moves and winning condition. Add code for winning condition.

@@ -220,13 +220,73 @@ class Engine:
                 elif self.board[pos] == 4:
                     line = line + bg_white + fg_black + "o" + bg_default + fg_default
             print(line)
-    def isWinning(self):
+    def is_winning_line(self, line):
+        '''Return who made a serie on this line.
+        -1: nobody
+         0: player 1
+         1: player 2
+         2: both
+        '''
+        victory = [False, False]
+        #Initializing with the first element of the line
+        prev_val = int(self.board[line[0]])
+        if prev_val != 0:
+            count = [1, 1]
+        else:
+            count = [0, 0]
+
+        #Start the loop over the rest of the line
+        for coor in line[1:]:
+            val = int(self.board[coor])
+            if val == 0:
+                count = [0, 0]
+            elif prev_val == 0:
+                count = [1, 1]
+            else:
+                #For both color and dot, check if the previous value match the current value so the serie keep going.
+                #For instance, if prev_val and val indicate the same color or the same type of dot.
+                for i in range(2):
+                    if self.matching_cells[i][prev_val-1][val-1]:
+                        count[i] += 1
+                    else:
+                        count[i] = 1
+                    #NOTE: Don't break if one is winning because it may be a double win
+                    victory[i] = victory[i] or (count[i] >= self.win_length)
+            prev_val = val
+
+        if victory[0] and victory[1]:
+            return 2
+        if victory[0]:
+            return 0
+        if victory[1]:
+            return 1
+        return -1
+    def is_winning(self):
         '''Return who is winning.
         -1: nobody
          0: player 1
          1: player 2
          2: tie
         '''
+        #TODO improve the search with a maximum high
+        victory = [False, False]
+        #Loop through all the lines. Don't break soon, because we need to check if it's a move with double victory
+        for line in self.all_lines:
+            val = self.is_winning_line(line)
+            if val != -1:
+                if val != 2:
+                    victory[val] = True
+                else:
+                    victory = [True, True]
+            if victory[0] and victory[1]:
+                return (len(self.previous_moves)-1)%2
+
+        if victory[0]:
+            return 0
+        elif victory[1]:
+            return 1
+        if len(self.previous_moves) > 60:
+            return 2
         return -1
     def play(self, player1, player2):
         self.ais = [player1, player2]
@@ -241,6 +301,6 @@ class Engine:
                 continue
             current_player = (current_player+1)%2
 
-            who_win = self.isWinning()
+            who_win = self.is_winning()
             if who_win >= 0:
                 return who_win

### 2019-01-26 tests/test_engine_move.py

@@ -143,7 +143,7 @@ def test_move_empty():
     assert(val == False)
 
 def test_move_right3():
-    #Add a card then a vertical card on top of it
+    '''Add a card then a vertical card on top of it'''
     engine = Engine()
     move = Move()
     move.recycling = False
@@ -157,6 +157,7 @@ def test_move_right3():
     assert(val == True)
 
 def test_move_right2():
+    '''Add a card then add another card somewhere else.'''
     engine = Engine()
     move = Move()
     move.recycling = False

### 2019-01-26 tests/test_engine_win.py

@@ -1,0 +1,148 @@
+from engine import Engine
+from move import Move
+
+def test_win_simple_color():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    engine.execute(move)
+    move.pos = (0,1)
+    engine.execute(move)
+    move.pos = (0,2)
+    engine.execute(move)
+    move.type = 8 # type 8 so only color win, with only one column
+    move.pos = (0,3)
+    engine.execute(move)
+
+    val = engine.is_winning()
+    #NOTE because color are winning and color is player 0
+    assert(val == 0)
+
+def test_win_simple_dot():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    engine.execute(move)
+    move.pos = (0,1)
+    engine.execute(move)
+    move.pos = (0,2)
+    engine.execute(move)
+    move.type = 6 # type 6 so only dot win, with only one column
+    move.pos = (0,3)
+    engine.execute(move)
+
+    val = engine.is_winning()
+    #NOTE because dot are winning and dot is player 1
+    assert(val == 1)
+
+def test_double_win():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    engine.execute(move) #Player 1 play
+    move.pos = (0,1)
+    engine.execute(move) #Player 2 play
+    move.pos = (0,2)
+    engine.execute(move) #Player 1 play
+    move.pos = (0,3)
+    engine.execute(move) #Player 2 play
+
+    #Because both player should win but player 2 was the last one to play, player 2 should win
+    val = engine.is_winning()
+    assert(val == 1)
+
+def test_win_diagonal_color():
+    engine = Engine()
+    move = Move()
+    move.recycling = False
+    move.type = 1
+    move.pos = (0,0)
+    engine.execute(move) #Player 1 play
+    move.pos = (2,0)
+    engine.execute(move) #Player 2 play
+    move.pos = (1,1)
+    engine.execute(move) #Player 1 play
+    move.pos = (4,0)
+    engine.execute(move) #Player 2 play
+    move.type = 6
+    move.pos = (3,1)
+    engine.execute(move) #Player 2 play
+    move.pos = (4,1)
+    engine.execute(move) #Player 2 play
[…]

### 2019-01-26 src/human_player.py
Add human player.

@@ -1,0 +1,57 @@
+import sys
+import re
+from engine import Engine
+from move import Move
+
+class HumanPlayer:
+    def __init__(self, name="otter"):
+        self.name = name
+        pattern = "^0 ([1-8]) ([A-H]|[a-h]) ([1-9]$|1[0-2]$)"
+        self.re_regular = re.compile(pattern)
+        pattern = "^([A-H]|[a-h]) ([1-9]$|1[0-2]$) ([A-H]|[a-h]) ([1-9]$|1[0-2]$) ([1-8]) ([A-H]|[a-h]) ([1-9]$|1[0-2]$)"
+        self.re_recycling = re.compile(pattern)
+
+    def play(self, engine):
+        engine.print()
+        while True:
+            print('(\033[33m' + self.name + '\033[00m) Move: ', end='', flush=True)
+            line = sys.stdin.readline()
+            matching = self.re_regular.match(line)
+            move = None
+            #Check if input match regular move
+            if matching is not None:
+                move = Move()
+                move.recycling = False
+                move.type = int(matching.group(1))
+                #Add two -1 because our coordinates start from zero
+                x = ord(matching.group(2)[0]) - 96 -1
+                y = int(matching.group(3)) - 1
+                move.pos = (x, y)
+            else:
+                #Otherwise, check if input match recycling move
+                matching = self.re_recycling.match(line)
+                if matching is not None:
+                    #Add -1 because our coordinates start from zero
+                    rec_x1 = ord(matching.group(1)[0]) - 96 -1
+                    rec_y1 = int(matching.group(2)) - 1
+                    rec_x2 = ord(matching.group(3)[0]) - 96 -1
+                    rec_y2 = int(matching.group(4)) - 1
+                    # We only use the bottom-left coordinate for the recycling
+                    rec_x = min(rec_x1, rec_x2)
+                    rec_y = min(rec_y1, rec_y2)
+                    t = int(matching.group(5))
+                    x = ord(matching.group(6)[0]) - 96 -1
+                    y = int(matching.group(7)) - 1
+                    move = Move()
+                    move.recycling = True
+                    move.type = t
+                    move.pos = (x, y)
+                    move.pos_rec = (rec_x, rec_y)
+            #If the move has been chosen, check if it's legal
+            if move is not None:
+                if engine.check_move(move):
+                    return move
+                else:
+                    print("Illegal move")
+            else:
+                print("Wrong input")

### 2019-01-26 tests/test_human_player.py
Add test for parsing human input.

@@ -1,0 +1,84 @@
+from human_player import HumanPlayer
+
+def test_input_regular():
+    player = HumanPlayer()
+    line = "0 1 c 8"
+    mv = player.match_move(line)
+    assert(mv is not None)
+    assert(mv.recycling == False)
+    assert(mv.type == 1)
+    assert(mv.pos[0] == 2) #We expect 2, because coordinates start at zero
+    assert(mv.pos[1] == 7) #Same here
+
+def test_input_regular2():
+    player = HumanPlayer()
+    line = "0 1 c 10"
+    mv = player.match_move(line)
+    assert(mv is not None)
+    assert(mv.recycling == False)
+    assert(mv.type == 1)
+    assert(mv.pos[0] == 2) #We expect 2, because coordinates start at zero
+    assert(mv.pos[1] == 9) #Same here
+
+def test_input_recycling():
+    player = HumanPlayer()
+    line = "f 3 F 4 7 c 8"
+    mv = player.match_move(line)
+    assert(mv is not None)
+    assert(mv.recycling == True)
+    assert(mv.pos_rec[0] == 5) #We expect 5, because coordinates start at zero
+    assert(mv.pos_rec[1] == 2) #Same here
+    assert(mv.type == 7)
+    assert(mv.pos[0] == 2) #We expect 2, because coordinates start at zero
+    assert(mv.pos[1] == 7) #Same here
+
+def test_input_wrong_type():
+    player = HumanPlayer()
+    line = "f 3 F 4 9 c 8"
+    mv = player.match_move(line)
+    assert(mv is None)
+    line = "f 3 F 4 0 c 8"
+    mv = player.match_move(line)
+    assert(mv is None)
+
+def test_input_wrong_regular():
+    player = HumanPlayer()
+    line = "0 1 c 01"
+    mv = player.match_move(line)
+    assert(mv is None)
+
+def test_input_negative():
+    player = HumanPlayer()
+    line = "0 1 c -1"
+    mv = player.match_move(line)
+    assert(mv is None)
+    line = "0 -1 c 1"
+    mv = player.match_move(line)
+    assert(mv is None)
+
+def test_input_no_adjacent_recycling():
+    player = HumanPlayer()
+    line = "f 3 F 5 1 c 8"
+    mv = player.match_move(line)
+    assert(mv is None)
+    line = "a 3 b 3 1 c 8"
+    mv = player.match_move(line)
[…]

### 2019-01-26 src/human_player.py
Fix HumanPlayer according to some tests that didn't pass.

@@ -8,45 +8,54 @@ class HumanPlayer:
         self.name = name
         pattern = "^0 ([1-8]) ([A-H]|[a-h]) ([1-9]$|1[0-2]$)"
         self.re_regular = re.compile(pattern)
-        pattern = "^([A-H]|[a-h]) ([1-9]$|1[0-2]$) ([A-H]|[a-h]) ([1-9]$|1[0-2]$) ([1-8]) ([A-H]|[a-h]) ([1-9]$|1[0-2]$)"
+        pattern = "^([A-H]|[a-h]) ([1-9]|1[0-2]) ([A-H]|[a-h]) ([1-9]|1[0-2]) ([1-8]) ([A-H]|[a-h]) ([1-9]$|1[0-2]$)"
         self.re_recycling = re.compile(pattern)
+
+    def match_move(self, line):
+        matching = self.re_regular.match(line)
+        move = None
+        #Check if input match regular move
+        if matching is not None:
+            move = Move()
+            move.recycling = False
+            move.type = int(matching.group(1))
+            #Add two -1 because our coordinates start from zero
+            x = ord(matching.group(2).lower()[0]) - 96 -1
+            y = int(matching.group(3)) - 1
+            move.pos = (x, y)
+        else:
+            #Otherwise, check if input match recycling move
+            matching = self.re_recycling.match(line)
+            if matching is not None:
+                #Add -1 because our coordinates start from zero
+                for i in range(8):
+                    print(matching.group(i))
+                rec_x1 = ord(matching.group(1).lower()[0]) - 96 -1
+                rec_y1 = int(matching.group(2)) - 1
+                rec_x2 = ord(matching.group(3).lower()[0]) - 96 -1
+                rec_y2 = int(matching.group(4)) - 1
+                # We only use the bottom-left coordinate for the recycling
+                dist = abs(rec_x1 - rec_x2) + abs(rec_y1 - rec_y2)
+                if dist != 1:
+                    return None
+                rec_x = min(rec_x1, rec_x2)
+                rec_y = min(rec_y1, rec_y2)
+                t = int(matching.group(5))
+                x = ord(matching.group(6).lower()[0]) - 96 -1
+                y = int(matching.group(7)) - 1
+                move = Move()
+                move.recycling = True
+                move.type = t
+                move.pos = (x, y)
+                move.pos_rec = (rec_x, rec_y)
+        return move
 
     def play(self, engine):
         engine.print()
         while True:
[…]

### 2019-01-26 src/move.py
Add fancy print for moves so it will be easier to understand what move it is.

@@ -4,3 +4,43 @@ class Move:
         self.type = 0
         self.pos = (-1,-1)
         self.pos_rec = (-1,-1)
+    def str_pos(self, pos):
+        return "[" + str(pos[0]) + ", " + str(pos[1]) + "]"
+    def __str__(self):
+        bg_default = "\033[49m"
+        bg_red = "\033[41m"
+        bg_white = "\033[107m"
+        fg_black = "\033[98m"
+        fg_default = "\033[00m"
+
+        if self.recycling:
+            line = "R " + self.str_pos(self.pos_rec) + " -> " + self.str_pos(self.pos)
+        else:
+            line = "M " + self.str_pos(self.pos)
+        line = line + " ("
+        #Create the upper line for type of move on 2 lines
+        if self.type in { 1, 3, 5, 7}:
+            upper_line = ""
+        else:
+            upper_line = "".join([" " for i in range(len(line))])
+        if self.type == 1:
+            line = line + bg_red + fg_black + "x" + bg_white + "o" + fg_default + bg_default + ")"
+        elif self.type == 2:
+            upper_line = upper_line + bg_red + fg_black + "x" + fg_default + bg_default + "\n"
+            line = line + bg_white + fg_black + "o" + fg_default + bg_default + ")"
+        elif self.type == 3:
+            line = line + bg_white + fg_black + "o" + bg_red + "x" + fg_default + bg_default + ")"
+        elif self.type == 4:
+            upper_line = upper_line + bg_white + fg_black + "o" + fg_default + bg_default + "\n"
+            line = line + bg_red + fg_black + "x" + fg_default + bg_default + ")"
+        elif self.type == 5:
+            line = line + bg_red + fg_black + "o" + bg_white + "x" + fg_default + bg_default + ")"
+        elif self.type == 6:
+            upper_line = upper_line + bg_red + fg_black + "o" + fg_default + bg_default + "\n"
+            line = line + bg_white + fg_black + "x" + fg_default + bg_default + ")"
+        elif self.type == 7:
+            line = line + bg_white + fg_black + "x" + bg_red + "o" + fg_default + bg_default + ")"
+        elif self.type == 8:
+            upper_line = upper_line + bg_white + fg_black + "x" + fg_default + bg_default + "\n"
+            line = line + bg_red + fg_black + "o" + fg_default + bg_default + ")"
+        return upper_line + line

### 2019-01-26 src/engine.py
Add function to get all possible regular moves.

@@ -187,6 +187,36 @@ class Engine:
                 return False
         #TODO: when recycling we should not undo the last move of our opponent
         return True
+    def available_moves(self):
+        if self.card_count == 0:
+            return self.available_recycling()
+        else:
+            return self.available_regular()
+    def available_recycling(self):
+        return []
+    def available_regular(self):
+        regular_moves = []
+        shape = self.board.shape
+        for x in range(shape[0]):
+            #Find the first no-empty cell of the column
+            pos = (x,0)
+            for y in range(shape[1]-1, -1, -1):
+                if self.board[(x,y)] != 0:
+                    pos = (x,y+1)
+            #If the pos we found is not even on the board, it means the column is full
+            if not self.is_on_board(pos):
+                continue
+            #For all type of position, we check if its a legal move and add it to the list if so.
+            for t in range(1, 9):
+                move = Move()
+                move.recycling = False
+                move.type = t
+                move.pos = pos
+                legal = self.check_move(move)
+                if legal:
+                    regular_moves.append(move)
+        return regular_moves
+
     def is_on_board(self, pos):
         if pos[0] < 0 or pos[0] > self.width:
             return False

### 2019-01-26 tests/test_engine_other.py
Add test functions to test out-of-board function.

@@ -1,0 +1,34 @@
+from engine import Engine
+
+def test_out_of_board():
+    engine = Engine(8, 12)
+    assert(engine.is_on_board((8, 0)) == False)
+    assert(engine.is_on_board((7, 0)) == True)
+    assert(engine.is_on_board((0, 0)) == True)
+    assert(engine.is_on_board((-1, 0)) == False)
+    assert(engine.is_on_board((7, 11)) == True)
+    assert(engine.is_on_board((7, 12)) == False)
+    assert(engine.is_on_board((8, 11)) == False)
+    assert(engine.is_on_board((8, 12)) == False)
+    assert(engine.is_on_board((3, 12)) == False)
+    assert(engine.is_on_board((3, -12)) == False)
+    assert(engine.is_on_board((3, 10)) == True)
+
+def test_out_of_board2():
+    engine = Engine(3, 3)
+    assert(engine.is_on_board((8, 0)) == False)
+    assert(engine.is_on_board((7, 0)) == False)
+    assert(engine.is_on_board((0, 0)) == True)
+    assert(engine.is_on_board((-1, 0)) == False)
+    assert(engine.is_on_board((7, 11)) == False)
+    assert(engine.is_on_board((7, 12)) == False)
+    assert(engine.is_on_board((8, 11)) == False)
+    assert(engine.is_on_board((8, 12)) == False)
+    assert(engine.is_on_board((3, 12)) == False)
+    assert(engine.is_on_board((3, -12)) == False)
+    assert(engine.is_on_board((3, 10)) == False)
+    assert(engine.is_on_board((2, 2)) == True)
+    assert(engine.is_on_board((3, 3)) == False)
+    assert(engine.is_on_board((0, 3)) == False)
+    assert(engine.is_on_board((3, 0)) == False)
+    assert(engine.is_on_board((2, 0)) == True)

### 2019-01-26 src/engine.py
Fix out-of-board issue.

@@ -212,15 +212,16 @@ class Engine:
                 move.recycling = False
                 move.type = t
                 move.pos = pos
+                print(move)
                 legal = self.check_move(move)
                 if legal:
                     regular_moves.append(move)
         return regular_moves
 
     def is_on_board(self, pos):
-        if pos[0] < 0 or pos[0] > self.width:
+        if pos[0] < 0 or pos[0] >= self.width:
             return False
-        if pos[1] < 0 or pos[1] > self.height:
+        if pos[1] < 0 or pos[1] >= self.height:
             return False
         return True
     def print(self):

### 2019-01-26 src/random_player.py
Fix random player to return a move.

@@ -1,0 +1,11 @@
+import numpy as np
+from engine import Engine
+
+class RandomPlayer:
+    def __init__(self):
+        self.name = "Random"
+
+    def play(self, engine):
+        all_moves = engine.available_moves()
+        idx = np.random.randint(0, len(all_moves))
+        return all_moves[idx]

### 2019-01-27 src/engine.py
Fix a bug in the engine.

@@ -169,17 +169,24 @@ class Engine:
         # placement_pos is where the card will be placed, thus, all position in this array should be empty and on the board.
         # recycling_pos contains the current position of the card to recycle
 
+        #Create the lambda to check the recycling position
+        if move.recycling:
+            l_in_rec = lambda p: np.any(np.all(recycling_pos == p, axis=1))
+        else:
+            l_in_rec = lambda p: False
+
+	#Check is placement_pos are available
         for i in range(placement_pos.shape[0]):
             pos = tuple(placement_pos[i])
             #If we recycle a card partially on itself, we need to consider the recycled cells as empty
-            if not self.is_on_board(pos) or (self.board[pos] != 0 and pos not in recycling_pos):
+            if not (self.is_on_board(pos) and (self.board[pos] == 0 or l_in_rec(pos))):
                 return False
         for i in range(no_empty_pos.shape[0]):
             pos = tuple(no_empty_pos[i])
             on_board = self.is_on_board(pos)
             is_empty = self.board[pos] == 0
             #If we recycle a card to put it above, we need to consider the cells recycled as empty
-            if on_board and (is_empty or pos in recycling_pos): #on board and empty
+            if on_board and (is_empty or l_in_rec(pos)): #on board and empty
                 return False
         for i in range(empty_pos.shape[0]):
             pos = tuple(empty_pos[i])
@@ -193,7 +200,56 @@ class Engine:
         else:
             return self.available_regular()
     def available_recycling(self):
-        return []
+        print("Recycling move!!!")
+        shape = self.board.shape
+        all_moves = []
+        for x in range(shape[0]):
+            for y in range(shape[1]-1, -1, -1):
+                pos  = (x,y)
+                val = self.cards[pos]
+                if val != 0 and pos != self.previous_moves[-1].pos:
+                    all_moves.extend(self.available_recycling_card(pos))
+                    break
+        return all_moves
+    def available_recycling_card(self, card_pos):
+        '''Return all the recycling move for one card located at card_pos'''
+        shape = self.board.shape
+        card_type = int(self.cards[card_pos])
+        ret = []
+        #Store the x corresponding to card_pos to retrieve them later
+        if card_type in {1, 3, 5, 7}: #If it is an horizontal move we need to x (thus two column are concerned).
+            card_pos_x = np.array([card_pos[0], card_pos[0]+1])
+        else:
+            card_pos_x = np.array([card_pos[0]])
+
[…]

### 2019-01-27 tests/test_engine_move.py

@@ -1,5 +1,6 @@
 from engine import Engine
 from move import Move
+from human_player import HumanPlayer
 
 def test_move_wrong_recycling():
     engine = Engine()
@@ -7,7 +8,7 @@ def test_move_wrong_recycling():
     move.recycling = True
     move.type = 1
     move.pos = (0,0)
-    val = engine.execute(move)
+    val = engine.check_move(move)
     assert(val == False)
 
 def test_recycling():
@@ -168,3 +169,22 @@ def test_move_right2():
     move.pos = (2,0)
     val = engine.execute(move)
     assert(val == True)
+
+def test_recycling_on_card():
+    engine = Engine(card_count=2)
+    human = HumanPlayer()
+    mv_str = "0 6 A 1"
+    mv = human.match_move(mv_str)
+    assert(mv is not None)
+    val = engine.execute(mv)
+    assert(val == True)
+    mv_str = "0 8 G 1"
+    mv = human.match_move(mv_str)
+    assert(mv is not None)
+    val = engine.execute(mv)
+    assert(val == True)
+    mv_str = "A 1 A 2 7 F 1"
+    mv = human.match_move(mv_str)
+    assert(mv is not None)
+    val = engine.execute(mv)
+    assert(val == False)

### 2019-01-27 src/move.py
Add function to output moves as input for HumanPlayer.

@@ -44,3 +44,17 @@ class Move:
             upper_line = upper_line + bg_white + fg_black + "x" + fg_default + bg_default + "\n"
             line = line + bg_red + fg_black + "o" + fg_default + bg_default + ")"
         return upper_line + line
+    def print_as_input(self):
+        line = ""
+        if self.recycling:
+            line += chr(self.pos_rec[0] + 65)
+            line += " " + str(self.pos_rec[1]+1)
+            #This is random because we don't the second position of the card only the bottom left is used so far
+            line += " " + chr(self.pos_rec[0] + 65)
+            line += " " + str(self.pos_rec[1]+2)
+        else:
+            line += "0"
+        line += " " + str(self.type)
+        line += " " + chr(self.pos[0] + 65)
+        line += " " + str(self.pos[1]+1)
+        print(line)

### 2019-01-27 src/human_player.py
Remove print in HumanPlayer.

@@ -28,8 +28,6 @@ class HumanPlayer:
             matching = self.re_recycling.match(line)
             if matching is not None:
                 #Add -1 because our coordinates start from zero
-                for i in range(8):
-                    print(matching.group(i))
                 rec_x1 = ord(matching.group(1).lower()[0]) - 96 -1
                 rec_y1 = int(matching.group(2)) - 1
                 rec_x2 = ord(matching.group(3).lower()[0]) - 96 -1

### 2019-01-27 src/engine.py
Remove some print in engine.

@@ -200,7 +200,6 @@ class Engine:
         else:
             return self.available_regular()
     def available_recycling(self):
-        print("Recycling move!!!")
         shape = self.board.shape
         all_moves = []
         for x in range(shape[0]):
@@ -384,9 +383,9 @@ class Engine:
             move = self.ais[current_player].play(self) #Call the player function
             # print("Player ", current_player, ": ")
             # print(move)
-            print("Turn: ", current_turn," ",self.card_count)
+            # print("Turn: ", current_turn," ",self.card_count)
             legal_move = self.execute(move)
-            self.print()
+            # self.print()
             if not legal_move:
                 continue
             current_player = (current_player+1)%2

### 2019-01-27 src/engine.py
Add a function to cancel move instead of deep copying the engine in minmax.

@@ -91,11 +91,57 @@ class Engine:
         if legal:
             self.do_move(move)
         return legal
+    def cancel_last_move(self):
+        last_move = self.previous_moves[-1]
+        self.previous_moves.pop()
+        #Recycling or not, remove the card movead
+        #Do it before the recycling in order to avoid erase the card replaced by a recycling
+        if last_move.type in {1, 3, 5, 7}: # horizontal last_move
+            pos1 = last_move.pos
+            pos2 = (last_move.pos[0]+1, last_move.pos[1])
+        elif last_move.type in {2, 4, 6, 8}: # vertical last_move
+            pos1 = last_move.pos
+            pos2 = (last_move.pos[0], last_move.pos[1]+1)
+        self.cards[pos1] = 0
+        self.board[pos1] = 0
+        self.board[pos2] = 0
+
+        #if last move was a recycling one, put back the card
+        if last_move.recycling:
+            #Find the two position
+            if last_move.type_rec in {1, 3, 5, 7}: # horizontal last_move
+                pos1 = last_move.pos_rec
+                pos2 = (last_move.pos_rec[0]+1, last_move.pos_rec[1])
+            elif last_move.type_rec in {2, 4, 6, 8}: # vertical last_move
+                pos1 = last_move.pos_rec
+                pos2 = (last_move.pos_rec[0], last_move.pos_rec[1]+1)
+
+            #Set the value on the board
+            if last_move.type_rec == 1 or last_move.type_rec == 4:
+                self.board[pos1] = 1
+                self.board[pos2] = 4
+            elif last_move.type_rec == 2 or last_move.type_rec == 3:
+                self.board[pos1] = 4
+                self.board[pos2] = 1
+            elif last_move.type_rec == 5 or last_move.type_rec == 8:
+                self.board[pos1] = 2
+                self.board[pos2] = 3
+            elif last_move.type_rec == 6 or last_move.type_rec == 7:
+                self.board[pos1] = 3
+                self.board[pos2] = 2
+
+            #Place the type in the cards matrix
+            self.cards[pos1] = last_move.type_rec
+        else:
+            self.card_count += 1
+
+
     def do_move(self, move):
         # Now we've checked the move, let's do it
         if move.recycling:
             #Don't use recycling_pos because I intend to separate the function in two later
             pos1 = move.pos_rec
+            move.type_rec = self.cards[move.pos_rec]
[…]

### 2019-01-27 src/min_max_player.py

@@ -1,0 +1,75 @@
+import numpy as np
+import time
+import copy
+from engine import Engine
+import cProfile
+
+def timeit(method):
+    def timed(*args, **kw):
+        ts = time.time()
+        result = method(*args, **kw)
+        te = time.time()
+        if 'log_time' in kw:
+            name = kw.get('log_name', method.__name__.upper())
+            kw['log_time'][name] = int((te - ts) * 1000)
+        else:
+            print ('%r  %2.2f ms' % (method.__name__, (te - ts) * 1000))
+        return result
+    return timed
+
+class MinMaxPlayer:
+    def __init__(self):
+        self.name = "MinMax"
+
+    @timeit
+    def play(self, id_ai, engine):
+        local_engine = copy.deepcopy(engine)
+        move = self.max(id_ai, engine, 3, -1000000, 1000000)
+        return move[1]
+
+    def heuristic(self, id_ai, engine):
+        winner = engine.is_winning()
+        if winner == id_ai:
+            return 10000
+        elif winner == (id_ai + 1)%2:
+            return -10000
+        elif winner == 2:
+            return -1000
+        return np.random.randint(1, 10)
+
+    def max(self, id_ai, engine, depth, alpha, beta):
+        current_val = self.heuristic(id_ai, engine)
+        if current_val >= 10000 or current_val <= -10000 or depth == 0:
+            return (current_val, None)
+        moves = engine.available_moves()
+        best_val = -100000
+        best_move = None
+        for move in moves:
+            engine.do_move(move) #No need to check if the move is legal it has already been done
+            value = self.min((id_ai%2)+1, engine, depth-1, alpha, beta)
+            engine.cancel_last_move()
+            value = value[0] #Min return tuple
+            if value > alpha:
+                alpha = value
+                best_move = move
+            if value >= beta:
+                return (value, move)
+        return (alpha, best_move)
+
+    def min(self, id_ai, engine, depth, alpha, beta):
+        current_val = self.heuristic(id_ai, engine)
+        if current_val >= 10000 or current_val <= -10000 or depth == 0:
+            return (current_val, None)
+        moves = engine.available_moves()
+        best_move = None
+        for move in moves:
+            engine.do_move(move) #No need to check if the move is legal it has already been done
+            value = self.max((id_ai%2)+1, engine, depth-1, alpha, beta)
+            engine.cancel_last_move()
+            value = value[0] #Min return tuple
+            if value < beta:
+                beta = value
[…]

### 2019-01-28 src/engine.py
Rewrite is_wining function with convolutions.
Start re-writing check_move to remove some for loops.

@@ -1,12 +1,14 @@
 import numpy as np
 from move import Move
+from scipy import signal
+from scipy import misc
 
 class Engine:
     def __init__(self, width=8, height=12, card_count=24, max_turn=60):
         self.width = width
         self.height = height
-        self.board = np.zeros((width, height))
-        self.cards = np.zeros((width, height))
+        self.board = np.zeros((width, height), dtype=np.int8)
+        self.cards = np.zeros((width, height), dtype=np.int8)
         self.card_count = card_count
         self.max_turn = max_turn
         self.previous_moves = []
@@ -34,6 +36,11 @@ class Engine:
         self.matching_cells[1][3][1] = 1
 
         self.all_lines = self.build_all_lines()
+
+        self.conv_column = np.ones((1, self.win_length))
+        self.conv_row = np.ones((self.win_length, 1))
+        self.conv_diag = np.eye(self.win_length, dtype=np.int8)
+        self.conv_cdiag = self.conv_diag[::-1] #Flip on the vertical axis
 
     def build_all_lines(self):
         '''This function build all lines of the board with coordinates of each cell. Row, column, and diagonals'''
@@ -135,7 +142,6 @@ class Engine:
         else:
             self.card_count += 1
 
-
     def do_move(self, move):
         # Now we've checked the move, let's do it
         if move.recycling:
@@ -221,23 +227,49 @@ class Engine:
         else:
             l_in_rec = lambda p: False
 
-	#Check is placement_pos are available
-        for i in range(placement_pos.shape[0]):
+        def aaa():
+            #Check is placement_pos are available
+            i = 0
             pos = tuple(placement_pos[i])
             #If we recycle a card partially on itself, we need to consider the recycled cells as empty
             if not (self.is_on_board(pos) and (self.board[pos] == 0 or l_in_rec(pos))):
                 return False
-        for i in range(no_empty_pos.shape[0]):
-            pos = tuple(no_empty_pos[i])
-            on_board = self.is_on_board(pos)
-            is_empty = self.board[pos] == 0
-            #If we recycle a card to put it above, we need to consider the cells recycled as empty
-            if on_board and (is_empty or l_in_rec(pos)): #on board and empty
+
+            i = 1
+            pos = tuple(placement_pos[i])
+            #If we recycle a card partially on itself, we need to consider the recycled cells as empty
[…]

### 2019-03-21 tests/test_engine_win.py
Add more winning test cases.

@@ -1,5 +1,11 @@
 from engine import Engine
 from move import Move
+from human_player import HumanPlayer
+
+def play_move(engine, mv_str):
+    player = HumanPlayer()
+    mv = player.match_move(mv_str)
+    engine.execute(mv)
 
 def test_win_simple_color():
     engine = Engine()
@@ -292,3 +298,189 @@ def test_win_counter_diagonal_dot_magic():
     #Dot should win
     val = engine.is_winning()
     assert(val == 1)
+
+def test_no_win_high():
+    engine = Engine()
+    play_move(engine, "0 2 A 1")
+    play_move(engine, "0 4 B 1")
+    play_move(engine, "0 2 C 1")
+    play_move(engine, "0 4 D 1")
+    play_move(engine, "0 2 A 3")
+    play_move(engine, "0 2 B 3")
+    play_move(engine, "0 4 C 3")
+    play_move(engine, "0 4 D 3")
+    play_move(engine, "0 2 A 5")
+    play_move(engine, "0 4 B 5")
+    play_move(engine, "0 2 C 5")
+    play_move(engine, "0 4 D 5")
+    val = engine.is_winning()
+    assert(val == -1)
+
+def test_no_win_high_magic():
+    engine = Engine(dark_magic=True)
+    play_move(engine, "0 2 A 1")
+    play_move(engine, "0 4 B 1")
+    play_move(engine, "0 2 C 1")
+    play_move(engine, "0 4 D 1")
+    play_move(engine, "0 2 A 3")
+    play_move(engine, "0 2 B 3")
+    play_move(engine, "0 4 C 3")
+    play_move(engine, "0 4 D 3")
+    play_move(engine, "0 2 A 5")
+    play_move(engine, "0 4 B 5")
+    play_move(engine, "0 2 C 5")
+    play_move(engine, "0 4 D 5")
+    val = engine.is_winning()
+    assert(val == -1)
+
+def test_win_high_magic():
+    engine = Engine(dark_magic=True)
+    play_move(engine, "0 2 A 1")
+    play_move(engine, "0 2 B 1")
+    play_move(engine, "0 2 C 1")
+    play_move(engine, "0 4 D 1")
+    play_move(engine, "0 2 A 3")
+    play_move(engine, "0 2 B 3")
+    play_move(engine, "0 2 C 3")
[…]

### 2019-03-21 lib/magic.c
Add powered is_winning with dark magic. I am below the 3 second mark with depth 4 on my computer!!!

@@ -346,7 +346,7 @@ inline void count_line(int const x_start,
 	int prev_offset[2] = {-1, -1};
 	int val_prev_poffset[2] = {-1, -1};
 	int count[2];
-#define safe_size_temp_line (12+8)
+    #define safe_size_temp_line (12+8)
 	int tmp_vspaces[2][safe_size_temp_line]; //12+8 is just to be sure we have enough space
 	for(int i = 0; i < 2; ++i)
 		for(int j = 0; j < safe_size_temp_line; ++j)
@@ -582,6 +582,128 @@ static PyObject* heuristic(PyObject *dummy, PyObject *args)
 	return PyInt_FromLong(values[0] - values[1]);
 }
 
+inline void winning_count(int const x_start,
+					   int const y_start,
+ 					   int const x_offset_mul,
+					   int const y_offset_mul,
+					   double const* board,
+					   npy_intp const* shape_board,
+					   int align[2][4])
+{
+	int offset = 0;
+
+	int prev_val = -1;
+	int prev_offset[2] = {-1, -1};
+	int val_prev_poffset[2] = {-1, -1};
+	int count[2];
+	do{	
+		//Compute the current x and y
+		int x = x_start + offset*x_offset_mul;
+		int y = y_start + offset*y_offset_mul;
+		//Check if this position is on the board
+		if(x < 0 || x >= shape_board[0] || y < 0 || y >= shape_board[1]){
+			break;
+		}
+		//Get the real index and the current value
+		int idx = x*shape_board[1] + y;
+		int val = board[idx];
+		/*First turn, we need to initialize count*/
+		if(prev_val == -1){
+			if(val != 0)
+				count[0] = count[1] = 1;
+			else
+				count[0] = count[1] = 0;
+		}
+		else{
+			if(val == 0){ //Value is empty
+				if(prev_val > 0){ //The previous value wasn't empty, so we need to count a line of size count[i]
+					for(int i = 0; i < 2; ++i){
+						if(count[i] > 0)
+							align[i][count[i]-1] += 1;
+					}
+				}
+				count[0] = count[1] = 0;
+			}
+			else if(prev_val == 0){ //Value is not empty but previous value was empty
+				count[0] = count[1] = 1;
+			}
+			else{
+				/*For both color and dot, check if the previous value match the current value so the serie keep going.*/
[…]

### 2019-03-21 src/engine.py

@@ -5,7 +5,7 @@ from scipy import misc
 import magic
 
 class Engine:
-    def __init__(self, width=8, height=12, card_count=24, max_turn=60, colors=[0,1], dark_magic=False):
+    def __init__(self, width=8, height=12, card_count=24, max_turn=40, colors=[0,1], dark_magic=False, verbose=True):
         self.width = width
         self.height = height
         self.board = np.zeros((width, height), dtype=np.int8)
@@ -15,6 +15,7 @@ class Engine:
         self.previous_moves = []
         self.win_length = 4
         self.colors = colors
+        self.verbose = verbose
         self.dark_magic = dark_magic
         #matching_cells describes how cells can form lines together.
         #for instance, a cell with 1 (red with filled dot) share the same color as 2 (red with empty dot)
@@ -449,6 +450,18 @@ class Engine:
          1: Dot
          2: tie
         '''
+        if self.dark_magic:
+            ret = magic.is_winning(self.board)
+            if ret == 0:
+                return 0
+            if ret == 1:
+                return 1
+            if ret == 2:
+                player_idx = (len(self.previous_moves)-1)%2
+                return self.colors[player_idx]
+            if len(self.previous_moves) > 60:
+                return 2
+            return -1
         #NOTE: To move later as attribute
         board_color = np.array([[0,0], [-1, -1], [-1, 1], [1, -1], [1, 1]])
         board_remaped = board_color[self.board]
@@ -497,6 +510,8 @@ class Engine:
         self.ais = [player1, player2]
 
     def play(self, player1, player2):
+        import cProfile
+        import pstats
         '''Return which player has won.
         -1: nobody
          0: Color
@@ -508,10 +523,10 @@ class Engine:
         current_turn = 0
         while True:
             move = self.ais[current_player].play(current_player, self) #Call the player function
-            print("Turn: ", current_turn," ",self.card_count, " -> player ", self.ais[current_player].name, "(", current_player+1, ")")
-            move.print_as_input()
+            if self.verbose:
+                print("Turn: ", current_turn," ( card:",self.card_count, ") -> player ", self.ais[current_player].name, "(", current_player+1, ")")
+                move.print_as_input()
             legal_move = self.execute(move)
-            self.printy()
             if not legal_move:
                 continue
 

### 2019-03-21 src/main.py
Add more description in parameter of main.py.

@@ -20,10 +20,10 @@ import magic
 parser = argparse.ArgumentParser(description='Run the Double Card game.')
 parser.add_argument("--width", type=int, default=8, required=False,help="Set the width of the board (default: 8)")
 parser.add_argument("--height", type=int, default=12, required=False,help="Set the height of the board (default: 12)")
-parser.add_argument("--p1", type=str, default=["human"], nargs="*", required=False,help="Set the first player AI (default: human, demo_mm, demo_ab)")
+parser.add_argument("--p1", type=str, default=["human"], nargs="*", required=False,help="Set the first player AI (default: human, demo_mm, demo_ab, minmax, mc, random)")
 parser.add_argument("--h1", type=str, default=["random"], nargs="*", required=False,\
         help="Set the heuristic for the first player (default: random)(possible values are: basic, random, convolution, neural, vspace, demo)")
-parser.add_argument("--p2", type=str, default=["human"], nargs="*", required=False,help="Set the second player AI (default: human, demo_mm, demo_ab)")
+parser.add_argument("--p2", type=str, default=["human"], nargs="*", required=False,help="Set the second player AI (default: human, demo_mm, demo_ab, minmax, mc, random)")
 parser.add_argument("--h2", type=str, default=["random"], nargs="*", required=False,\
         help="Set the heuristic for the second player (default: random)(possible values are: basic, random, convolution, neural, vspace)")
 parser.add_argument("--run-type", type=str, default="standard", required=False,help="Indicate what to do (default: standard)(possible values: standard, win_rate, train)")
@@ -71,6 +71,7 @@ def load_vspace_heuristic(dm):
 
 #define the run functions
 def run_standard(args):
+    print(args)
     colors = [1, 0] if args.invert_colors else [0, 1]
     engine = Engine(args.width, args.height, colors=colors, dark_magic=args.dark_magic)
     winner = engine.play(args.p1, args.p2)

### 2019-03-21 lib/magic.c
Fix a stack crash in modules because lines size (of color or dot) could exceed 4 :].

@@ -337,7 +337,7 @@ inline void count_line(int const x_start,
 					   int const y_offset_mul,
 					   double const* board,
 					   npy_intp const* shape_board,
-					   int align[2][4],
+					   int align[2][10],
 					   int valuable_space[2][96])
 {
 	int offset = 0;
@@ -477,7 +477,7 @@ static PyObject* heuristic(PyObject *dummy, PyObject *args)
 		for(int j = 0; j < shape_board[0] * shape_board[1]; ++j)
 			valuable_space[i][j] = 0;
 	
-	int align[2][4]; //Contains the distribution of alignment length
+	int align[2][10]; //Contains the distribution of alignment length
 	for(int i = 0; i < 2; ++i)
 		for(int j = 0; j < 4; ++j)
 			align[i][j] = 0;
@@ -588,7 +588,7 @@ inline void winning_count(int const x_start,
 					   int const y_offset_mul,
 					   double const* board,
 					   npy_intp const* shape_board,
-					   int align[2][4])
+					   int align[2][10])
 {
 	int offset = 0;
 
@@ -668,9 +668,9 @@ static PyObject* is_winning(PyObject *dummy, PyObject *args)
 	npy_intp *shape_board = PyArray_SHAPE(npy_board);
 
 	double const* board = PyArray_DATA(npy_board);
-	int align[2][4]; //Contains the distribution of alignment length
+	int align[2][10]; //Contains the distribution of alignment length
 	for(int i = 0; i < 2; ++i)
-		for(int j = 0; j < 4; ++j)
+		for(int j = 0; j < 10; ++j)
 			align[i][j] = 0;
 	for(int col = 0; col < shape_board[0]; ++col){
 		//Check column from (col, 0)

### 2019-03-21 src/min_max_player.py
Add a timing check in MinMaxPlayer so it does not exceed 6 seconds even if it is not finished, yet. The best move found so far is returned.

@@ -35,15 +35,15 @@ class MinMaxPlayer:
         self.average_number_of_moves = 0
         self.number_of_state = 0
         self.heuristic_called_count = 0
-        move = self.max(id_ai, engine, self.depth, -100000000, 100000000)
-        # if len(engine.previous_moves) > 0:
-            # move = self.max(id_ai, engine, self.depth, -100000000, 100000000)
-        # else:
-            # moves = engine.available_moves()
-            #Exploring some first moves
-            # moves = [mv for mv in moves if (mv.pos[0] != 0 and mv.pos[0] != engine.width-1) and mv.type in {2, 4, 6, 8}]
-            # i = np.random.randint(0, len(moves))
-            # return moves[i]
+        self.stop_time = time.time()+5.97
+        if len(engine.previous_moves) > 0:
+            move = self.max(id_ai, engine, self.depth, -100000000, 100000000)
+        else:
+            moves = engine.available_moves()
+            # Exploring some first moves
+            moves = [mv for mv in moves if (mv.pos[0] != 0 and mv.pos[0] != engine.width-1) and mv.type in {2, 4, 6, 8}]
+            i = np.random.randint(0, len(moves))
+            return moves[i]
 
         if isinstance(move[1], list):
             for mv in move[1]:
@@ -110,6 +110,9 @@ class MinMaxPlayer:
             if value >= beta:
                 return (value, move)
                 # return (value, l_move)
+            if time.time() > self.stop_time:
+                return (alpha, best_move)
+
         return (alpha, best_move)
         # return (alpha, l_move)
 
@@ -150,5 +153,7 @@ class MinMaxPlayer:
             if value <= alpha:
                 return (value, move)
                 # return (value, l_move)
+            if time.time() > self.stop_time:
+                return (beta, best_move)
         return (beta, best_move)
         # return (beta, l_move)

### 2019-03-21 src/practice.py
Add moves to check a recycling situation with a high branching factor.

@@ -129,18 +129,49 @@ def genetic():
    print(hof)
 
 # genetic()
-# engine = Engine(dark_magic=True)
-# h = HeuristicVspace()
-# p = MinMaxPlayer(h, 3, False)
-# p2 = MinMaxPlayer(h)
-# engine.initialize_player(p2, p)
-# # play_move(engine, "0 6 F 1")
+engine = Engine(dark_magic=True)
+h = HeuristicVspace()
+p = MinMaxPlayer(h, 4, False)
+p2 = MinMaxPlayer(h)
+engine.initialize_player(p2, p)
+play_move(engine, "0 5 B 1")
+play_move(engine, "0 4 A 1")
+play_move(engine, "0 5 B 2")
+play_move(engine, "0 2 C 3")
+play_move(engine, "0 2 B 3")
+play_move(engine, "0 2 A 3")
+play_move(engine, "0 5 B 5")
+play_move(engine, "0 3 B 6")
+play_move(engine, "0 1 B 7")
+play_move(engine, "0 8 A 5")
+play_move(engine, "0 5 F 1")
+play_move(engine, "0 8 A 7")
+play_move(engine, "0 4 G 2")
+play_move(engine, "0 4 C 8")
+play_move(engine, "0 6 A 9")
+play_move(engine, "0 2 B 8")
+play_move(engine, "0 6 A 11")
+play_move(engine, "0 6 C 10")
+play_move(engine, "0 2 B 10")
+play_move(engine, "0 1 B 12")
+play_move(engine, "0 4 G 4")
+play_move(engine, "0 8 G 6")
+play_move(engine, "0 6 F 2")
+play_move(engine, "0 4 F 4")
+play_move(engine, "A 11 A 12 6 E 1")
+play_move(engine, "A 9 A 10 8 E 3")
+play_move(engine, "A 7 A 8 6 A 7")
+play_move(engine, "B 12 C 12 8 A 9")
+play_move(engine, "E 3 E 4 2 E 3")
+play_move(engine, "B 10 B 11 8 E 5")
+# play_move(engine, "0 6 F 1")
 
-# cProfile.run('mv = p.play(0, engine)', "output_stat")
-# p = pstats.Stats('output_stat')
-# p.strip_dirs()
-# p.sort_stats('tottime')
-# p.print_stats()
+cProfile.run('mv = p.play(0, engine)', "output_stat")
+print(mv)
+p = pstats.Stats('output_stat')
[…]

### 2019-03-21 report/paper.tex
Fill the report.

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usetikzlibrary{matrix}
 % Used for displaying a sample figure. If possible, figure files should
@@
 %
-%\authorrunning{COMP 472 [or COMP 6721] F. Author et al.}
 \institute{40093517 \email{martin.khannouz@gmail.com}}
@@
 %
-\section{First Section}
-\subsection{A Subsection Sample}
+\section*{Introduction}
+This project report describes how the first
+project of \textit{Introduction to Artificial
+Intelligence} were developed. The first section
+covers the environment and the tools used for the
+project. The heuristic section describes each
+heuristic and how they were improved. The
+difficulties section summerizes the difficulties
+encountered during the project and how they were
+overcome.
+
+\section{Settings and Structures}
+\subsection{Settings}
+The developing part of this project was realized
+on an Archlinux distribution with Python~3.7.2.
+Vim with a Jedi plugin was used for the coding
+part and a private Github repository was set to
+keep track of the changes.
+The \textit{orwell} machine was used before each
+deliverable to ensure the project worked with the
+software version on Concordia's computers. Pytest
+was used to guarantee no regression has been made.
+
+\subsection{stucture}
+The project is structured in classes with
+the main one being the engine. This class is an
+instance of the game that implements the rules.
+To ensure this class operates properly, unit tests
+were implemented before the class methods, based on
+the use cases given by the teacher.
+This class was optimized as much as possible to
+guarantee the fastest execution time for minimax
+player.
+
+The project also contains three type of players. A
+\textit{Human} player that parse the input from the command
+line. A \textit{MinMax} player that implement tweaked
+minimax version with alpha-beta pruning. The
+\textit{MonteCarlo} player which was supposed to implement
+a Monte Carlo Tree Search even though it was a
+complete failure due to the poor understanding of
+this algorithm by the author of the project. Finally, a random player was
+developed during the early stage of the project.
+This "algorithm" served as control test for other
+player types, because none of them should lost
+against such chaotic player\footnote{Based on my
+statistics, even a 5-years old kid should get a
+win rate of 50\%.}.
+
+Automatic players (\textit{MinMax} and
+\textit{MonteCarlo}) rely on heuristic to complete
+their search. All heuristic functions are grouped
+in one file: \textit{heuristic.py}.
+Finally, the class \textit{Move} represents a move
+that could be played.
+
+In addition to these classes, additional functions
+were implemented. One function used genetic
+algorithms to optimize the \textit{Vspace}
+heuristic, but it was too slow and just gave hint
+on how to enhance the heuristic rather than a
+accurate weights.  A win rate function was also
+implemented to compare combinations of player and
+heuristic against each other.
+
+\section{Heuristics}
+Five heuristics were developed in this project
+even though only two were made for competitive
+use. One heuristic is the naive function for the
+second deliverable. Another is a random heuristic
+that systematicaly return a new random number. It
+was used to spot runtime errors in \textit{MinMax} and
+\textit{MonteCarlo} algorithms. The \textit{Basic}
+heuristics only check for winning situation and
+was used for basic controls.
+The two other heuristics are described in
+subsection of their own.
+
+\subsection{Convolution}
+The convolution utilize the \textit{convolve2d} function
+from Scipy to estimate the value of a current
+state.
+The game board is split into two matrices: dot and
+color. Each matrix contains -1, 0, or 1. A zero
+indicates an empty space while the two other values
+represent either a filled dot and an empty dot, or
+a white cell and a red cell.
+Applying \textit{convolve2d} on these matrices
+produces new matrices with values that range from
+-4 to 4. An absolute value of four indicates a row
+of four and therefore, a winning position.
+The heuristic gathers all intermediate values into
+an histogram (using the \textit{histogram}
+function of numpy) then give a weight to each
+value.
+
+Figure~\ref{fig:board-example} shows an example of
+the board. Figure~\ref{fig:dot-representation}
+focus on the dot player and show the
+transformation applied to the board for this
+player. Filled dots are set to 1 and circles to
+-1. Zeros are not represented for more clarity.
+The heuristic applies four convolutions: one for
+the rows, one for the columns and one for each
+diagonal. These convolution generate four smaller
+matrices showed in
+Figure~\ref{fig:convolution-results}.  Then, the
+matrices are flatened, concateneted and the
+absolute value is applied to every cell. Finally
+the histogram is built out of this matrix and each
+element is multiplied by a weight before being
+summed.  The final
+value of the board is the difference between the
+two player values.
+
+This approach is strong for many reason. It is
+faster than calling python code because Scipy and
+Numpy rely on C writen functions. Uncomplete rows
+that contain zeros are more valued than uncomplete
+rows that contain opposite values. 
+\begin{figure}[ht]
+		\center
+		\begin{tikzpicture}
+
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {$\bullet$};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};
+
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {$\bullet$};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};
+
+		\end{tikzpicture}
+		\caption{An example of a board game. Cards are not represented.}
+		\label{fig:board-example}
+\end{figure}
+\begin{figure}[ht]
+		\center
+		\begin{tikzpicture}
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (A) at (2.0, 1) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0.5) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0.5) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0.5) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0.5) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (B) at (5, 1) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 1.5) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1.5) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1.5) {};
+
+				\draw [->,thick] (A) -- (B);
+		\end{tikzpicture}
+		\caption{Example of transformation for the dot player based on the board in Figure~\ref{fig:board-example}.}
+		\label{fig:dot-representation}
+\end{figure}
+\begin{figure}[ht]
+		\center
+		\begin{tikzpicture}
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {-2};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {-1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {2};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.5, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (3.0, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (C1) at (3.5, 0) {0};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1) {0};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (3.0, 1) {2};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (C0) at (3.5, 1) {-1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (H0) at (7.0, 0) {8};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.5, 0) {6};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (8.0, 0) {3};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (8.5, 0) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (9.0, 0) {0};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W0) at (7.0, 0.75) {0};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W1) at (7.5, 0.75) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W2) at (8.0, 0.75) {8};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W3) at (8.53, 0.75) {16};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W4) at (9.15, 0.75) {$10^6$};
+				\node at (10.5, 0.75) {Weights};
+				\node at (10.5, 0) {Histogram};
+				\node at (10.5, -0.5) {Indexes};
+
+				\node at (7.0, -0.5) {0};
+				\node at (7.5, -0.5) {1};
+				\node at (8.0, -0.5) {2};
+				\node at (8.5, -0.5) {3};
+				\node at (9.0, -0.5) {4};
+
+				\node (ID0) at (7.0, 2) {$\underbrace{0\times8} + $};
+				\node (ID1) at (8.0, 2) {$\underbrace{1\times6} +$};
+				\node (ID2) at (9.0, 2) {$\underbrace{8\times3} +$};
+				\node (ID3) at (10.15, 2) {$\underbrace{16\times0} +$};
+				\node (ID4) at (11.35, 2) {$\underbrace{10^6 \times 0}$};
+				\node (ID5) at (12.5, 2.10) {$ = 30$};
+
+				\draw [->,thick] (W0) -- (ID0);
+				\draw [->,thick] (W1) -- (ID1);
+				\draw [->,thick] (W2) -- (ID2);
+				\draw [->,thick] (W3) -- (ID3);
+				\draw [->,thick] (W4) -- (ID4);
+
+				\draw [->,thick] (C0) -- (H0);
+				\draw [->,thick] (C1) -- (H0);
+		\end{tikzpicture}
+		\caption{The four matrices obtained after applying the four masks on the matrix that represent the dot, in Figure~\ref{fig:dot-representation}.
+						This Figure also represents the conversion to an histogram and how this histogram is used to find the heuristic value for the dot player.}
+		\label{fig:convolution-results}
+\end{figure}
+
+\subsection{Vspace}
+The Vspace heuristic tries to give value to empty
+cells in addition to counting rows of similar
+symbols. This heuristic count the number of lines
+and their sizes.
+\section{Difficulties}
+\begin{itemize}
+		\item Slowness of python, (numpy first)
+		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)
+\end{itemize}
+\section{Analyze}
+\begin{itemize}
+		\item How is the pruning.
+		\item How are the performance at the tournament.
+\end{itemize}
+\section{A Subsection Sample}
 Please note that the first paragraph of a section or subsection is

### 2019-03-25 report/paper.tex
Add part for vspace.

@@
 project of \textit{Introduction to Artificial
-Intelligence} were developed. The first section
+Intelligence} was developed. The first section
 covers the environment and the tools used for the
@@
 on an Archlinux distribution with Python~3.7.2.
-Vim with a Jedi plugin was used for the coding
-part and a private Github repository was set to
+Vim with the Jedi plugin was used to code and a private Github repository was set to
 keep track of the changes.
@@
 The project is structured in classes with
-the main one being the engine. This class is an
-instance of the game that implements the rules.
+the main one representating the engine. This class is an
+instance of the game and implements the rules.
 To ensure this class operates properly, unit tests
@@
 guarantee the fastest execution time for minimax
-player.
+algorithm.
 
@@
 implemented to compare combinations of player and
-heuristic against each other.
+heuristic against each other. Finally, a non
+tweaked MinMax version was developed for
+deliverable 2.
 
@@
 produces new matrices with values that range from
--4 to 4. An absolute value of four indicates a row
+-4 to 4. A cell with an absolute value of four indicates a row
 of four and therefore, a winning position.
@@
 function of numpy) then give a weight to each
-value.
+value. For instance, a weight of 16 was given to
+each intermediate value of 3 which represent a row
+of three.
 
@@
 The heuristic applies four convolutions: one for
-the rows, one for the columns and one for each
+the rows, another for the columns and one for each
 diagonal. These convolution generate four smaller
 matrices showed in
-Figure~\ref{fig:convolution-results}.  Then, the
-matrices are flatened, concateneted and the
+Figure~\ref{fig:convolution-results} (left side).
+The matrices are then flatened, concateneted and the
 absolute value is applied to every cell. Finally
@@
 that contain zeros are more valued than uncomplete
-rows that contain opposite values. 
+rows that contain opposite values. Finally,
+because the convolution mask is applied to empty
+spaces, the heuristic can build complexe
+structures with space in the middle or above.
+
 \begin{figure}[ht]
@@
 
+\clearpage
 \subsection{Vspace}
-The Vspace heuristic tries to give value to empty
+The goal of the Vspace heuristic is to give value to empty
 cells in addition to counting rows of similar
-symbols. This heuristic count the number of lines
-and their sizes.
+symbols. Similarly to the previous heuristic,
+Vspace count the size of the lines, but for every
+line connected to an empty cell, it uses the
+size of this line as a value for the empty cell.
+
+Figure~\ref{fig:dot-vspace} and
+Figure~\ref{fig:color-vspace} give examples on
+how the spaces are valued for both players, based
+on the board at Figure~\ref{fig:board-example}.
+Figure~\ref{fig:color-vspace} shows that even if
+the line is split in two with an empty cell in the
+middle, the combine value of both side gives the
+value of the space. This is the reason why the two
+3-values appears even though there is no line of
+three.
+
+Possible value for spaces range from 0 to 4 and
+higher values are downed to 4. When multiple lines
+are connected to an empty cell, the cell receives
+the maximum value from them. The heuristic split the
+space in two categories: the space available in
+the next turn and the spaces too high to be reach
+with one card.
+
+Finally, an histogram for each category is made and
+each value is multiplied by a weight then summed.
+Note that the two categories have different
+weights, because a space connected to a line of
+three symboles does not hold the same value if you
+can play on it on the next turn or not.
+\begin{figure}[ht]
+		\center
+		\begin{tikzpicture}
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {$\circ$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {$\bullet$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {$\bullet$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {$\circ$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {$\circ$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] (A) at (2.0, 1) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {$\bullet$};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};
+
+				%------
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0) {$\circ$};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0) {$\bullet$};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0.5) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0.5) {$\bullet$};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0.5) {1};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0.5) {$\circ$};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0.5) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (B) at (5, 1) {$\circ$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1) {$\circ$};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1) {2};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1) {1};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 1.5) {$\bullet$};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1.5) {$\bullet$};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1.5) {2};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1.5) {};
+
+				\draw [->,thick] (A) -- (B);
+		\end{tikzpicture}
+		\caption{Example of the Vspace process for the
+		dot player based on the board in
+		Figure~\ref{fig:board-example}. the grey cells
+		represent cells that contain nothing
+		interesting. On the right board, green cells
+		are spaces available for the next turn and
+		purple cells are spaces that need at least two
+		cards to be filled.}
+		\label{fig:dot-vspace}
+\end{figure}
+\begin{figure}[ht]
+		\center
+		\begin{tikzpicture}
+
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};
+
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1) {};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};
+
+				%------
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (5, 0) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0) {};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0) {3};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0) {};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0.5) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0.5) {};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0.5) {3};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0.5) {};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0.5) {1};
+
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] (B) at (5, 1) {};
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1) {};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1) {1};
+				\node [draw, fill=green!15, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1) {1};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1) {1};
+
+				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 1.5) {};
+				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1.5) {$\bullet$};
+				\node [draw, fill=purple!25, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1.5) {2};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1.5) {};
+				\node [draw, fill=black!20, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1.5) {};
+
+				\draw [->,thick] (A) -- (B);
+		\end{tikzpicture}
+		\caption{Example of the Vspace process for the
+		color player based on the board in
+		Figure~\ref{fig:board-example}. the grey cells
+		represent cells that contain nothing
+		interesting. On the right board, green cells
+		are spaces available for the next turn and
+		purple cells are spaces that need at least two
+		cards to be filled.}
+		\label{fig:color-vspace}
+\end{figure}
+\clearpage
 \section{Difficulties}
@@
 \begin{itemize}
-		\item How is the pruning.
+		\item How is the pruning depth 3 with a
+				recycling moves: no prun = 26s, prun=0.3s.
 		\item How are the performance at the tournament.
 \end{itemize}
+
 \section{A Subsection Sample}

### 2019-03-27 report/paper.tex
Add report.

@@
 project. The heuristic section describes each
+%TODO Briefly describe what a heuristic is.
 heuristic and how they were improved. The
@@
 
+\subsection{Tweaks in \textit{MinMax}}
+The original \textit{MinMax} algorithm explore the
+complete tree of possibilities and applies the
+heuristic function on the leaf. However, the
+heuristic functions described below do not take in
+account who won first. For instance, a sequence of
+move could be prefered if a player scores three
+line of four even if his opponent already won at
+the previous move with only one line of four.
+
+To prevent this behavior, the MinMax algorithm was
+tweaked to check for winning condition after
+playing each possible move at any depth. This
+modification prevent the exploration of game
+states already won by a player.
+
+Even though calling the winning function from the
+engine was costly at the beginning of the project,
+I managed to drastically reduce this time by
+improving the performances.
+
 \section{Heuristics}
+%TODO add citation to wikipedia for heuristics
+%TODO add importance of the heuristic in minmax
 Five heuristics were developed in this project
@@
 \subsection{Vspace}
+\label{sec:vspace}
 The goal of the Vspace heuristic is to give value to empty
@@
 \section{Difficulties}
+\subsection{Slowness}
+The biggest difficulty encountered during this
+project was the slowness of python. In the early
+stage of the project, the code was
+too ineficient and could not check more than one
+move in advance within the limit of three seconds.
+
+The \textit{cProfile} python module is a module
+that helps to profile python code. In particular,
+it give the amount of time spent in each function
+called. This profiler was used to locate the most
+time consumming function, then I rely on this information to
+restructure the code into better object structure.
+
+However, the result were hardly perceivable and
+the code was still too slow for competetive use.
+Therefore, I turn to two scientific python
+libraries: Numpy and Scipy. Using functions from
+these libraries slightly improved the performance
+because they were built-in functions. In python,
+built-in functions are C functions compiled into a
+python module and called from python script.
+
+Finally, to reach a decent executing time, 
+one more step was taken and a python module writen
+in C was developed.
+Figure~\ref{fig:performance-magic} show the
+evolution of the performances as the most time
+consuming function were adapted to call the
+module. However, the python code remains and it is
+executed by default. 
+Figure~\ref{fig:performance-magic} shows that the
+MinMax algorithm is able to check four moves in
+advance after half of the key part used the
+module.
+
+\begin{figure}
+		\includegraphics[width=\textwidth]{perfromances.png}
+		\caption{The amount of time (in second) taken
+		to run the MinMax algorithm on a board with the
+		card "5 B 1" already played. The MinMax
+		algorithm checked up to 4 moves in advance and
+		it used the Vspace heuristic described in
+		Section~\ref{sec:vspace}. The left column
+		shows the full python execution time (more
+		than 4 minutes) and the other columns show the
+		execution time as part of the code start using
+		the C module. The third column from the left,
+		"checking a move", indicate the time taken
+		when the heuristic function and the function
+		that check the moves use both the C module.}
+		\label{fig:performance-magic}
+\end{figure}
+
 \begin{itemize}
-		\item Slowness of python, (numpy first)
 		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)

### 2019-03-27 src/heuristics.py
MuggleVspace.

@@ -95,9 +95,158 @@ class HeuristicMuggleVspace():
         self.name = "vspace"
         self.weights = np.array(weights)
 
+
+    def count_line(self, x_start, y_start, x_offset_mul, y_offset_mul, board, shape_board, align, valuable_space, engine):
+        offset = 0
+        prev_val = -1
+        prev_offset = [-1, -1]
+        val_prev_poffset = [-1, -1]
+        count = [0, 0]
+        tmp_vspaces = [[0 for i in range(20)] for j in range(2)]
+
+        #start do while
+        while 1:
+            x = x_start + offset*x_offset_mul
+            y = y_start + offset*y_offset_mul
+            if x < 0 or x >= shape_board[0] or y < 0 or y >= shape_board[1]:
+                break
+            idx = x*shape_board[1] + y
+            val = board[idx]
+            if prev_val == -1:
+                if val != 0:
+                    count[0] = count[1] = 1
+                else:
+                    count[0] = count[1] = 0
+                    prev_offset[0] = prev_offset[1] = offset
+            else:
+                if val == 0: #Value is empty
+                    if prev_val > 0: #The previous value wasn't empty, so we need to count a line of size count[i]
+                        for i in range(2):
+                            if count[i] > 0:
+                                align[i][count[i]-1] += 1
+                                tmp_vspaces[i][offset] += count[i]
+
+                                if prev_offset[i] >= 0:
+                                    po = prev_offset[i]
+                                    v = val_prev_poffset[i]
+                                    is_matching = engine.matching_cells[i][prev_val-1][v-1]
+                                    if v > 0 and is_matching: #If the value before the previous space match our current line, combine the vspace
+                                        tmp_vspaces[i][po] += count[i]
+                                    else:
+                                        tmp_vspaces[i][po] = max(count[i], tmp_vspaces[i][po])
+                    count[0] = count[1] = 0
+                    prev_offset[0] = prev_offset[1] = offset
+                    val_prev_poffset[0] = val_prev_poffset[1] = prev_val
+                elif prev_val == 0: #Value is not empty but previous value was empty
+                        count[0] = count[1] = 1
+                else:
+                    #For both color and dot, check if the previous value match the current value so the serie keep going.
+                    #For instance, if prev_val and val indicate the same color or the same type of dot.
+                    for i in range(2):
[…]

### 2019-03-28 report/paper.tex
Report.

--- 2019-03-25 report/paper.tex
@@
 \usepackage{tikz}
+\usepackage{subcaption}
+\captionsetup{compatibility=false}
 \usetikzlibrary{matrix}
@@
 %
-\title{COMP 6721: Project Report}
+\title{COMP 6721: Double Card}
 \author{Martin Khannouz}
@@
 project. The heuristic section describes each
+%TODO Briefly describe what a heuristic is.
 heuristic and how they were improved. The
-difficulties section summerizes the difficulties
+difficulties section summarizes the difficulties
 encountered during the project and how they were
@@
 
-\subsection{stucture}
+\subsection{Stucture}
 The project is structured in classes with
-the main one representating the engine. This class is an
+the main one representing the engine. This class is an
 instance of the game and implements the rules.
@@
 This class was optimized as much as possible to
-guarantee the fastest execution time for minimax
-algorithm.
+guarantee the fastest execution time for
+\textit{MinMax} algorithm.
 
-The project also contains three type of players. A
+The project also contains three types of players. A
 \textit{Human} player that parse the input from the command
@@
 a Monte Carlo Tree Search even though it was a
-complete failure due to the poor understanding of
-this algorithm by the author of the project. Finally, a random player was
+complete failure due to my poor understanding of
+this algorithm. Finally, a random player was
 developed during the early stage of the project.
-This "algorithm" served as control test for other
-player types, because none of them should lost
-against such chaotic player\footnote{Based on my
+This "algorithm" served as a control test for other
+player types because none of them should lose
+against such a chaotic player\footnote{Based on my
 statistics, even a 5-years old kid should get a
@@
 algorithms to optimize the \textit{Vspace}
-heuristic, but it was too slow and just gave hint
-on how to enhance the heuristic rather than a
+heuristic, but it was too slow and just gave a hint
+of how to enhance the heuristic rather than
 accurate weights.  A win rate function was also
 implemented to compare combinations of player and
-heuristic against each other. Finally, a non
-tweaked MinMax version was developed for
+heuristic against each other. Finally, a
+non-tweaked MinMax version was developed for
 deliverable 2.
 
+\subsection{Tweaks in \textit{MinMax}}
+The original \textit{MinMax} algorithm explores the
+complete tree of possibilities and applies the
+heuristic function on the leaf. However, the
+heuristic functions described below do not take
+into account who won first. For instance, a sequence of
+move could be preferred if a player scores three
+lines of four symbols even if his opponent already won at
+the previous move.
+
+To prevent this behavior, the MinMax algorithm was
+tweaked to check for winning condition after
+playing each possible move at any depth. This
+modification prevents the exploration of game
+states already won by a player.
+
+Even though calling the winning function from the
+engine was costly at the beginning of the project,
+I managed to drastically reduce this time by
+improving the performances.
+
+To prevent exceeding the six-second limit, a timer
+was added in the MinMax algorithm. If this limit
+was reached, the algorithm stops and returns the
+best move encountered until this time.
+
+The \textit{MinMax} algorithm contains a parameter
+to sort the moves at each level. This approach was
+supposed to prune the state space faster. At the
+beginning of the project, this sort improved
+significantly the performance, but on later
+versions of the heuristics, it slowed the algorithm.
+Therefore, it was barely used nor study.
+
+Finally, when the algorithm runs on an empty
+board, it returns a random vertical move. A
+vertical move because the heuristics appear to
+have a better winning rate when they start with
+this move type. The move was randomly picked to
+explore openings and to bring small changes from one game to another.
 \section{Heuristics}
+%TODO add citation to wikipedia for heuristics
+%TODO add importance of the heuristic in minmax
 Five heuristics were developed in this project
@@
 second deliverable. Another is a random heuristic
-that systematicaly return a new random number. It
+that systematically returns a new random number. It
 was used to spot runtime errors in \textit{MinMax} and
 \textit{MonteCarlo} algorithms. The \textit{Basic}
-heuristics only check for winning situation and
+heuristics only check for winning situations and
 was used for basic controls.
 The two other heuristics are described in
-subsection of their own.
+subsections of their own.
 
 \subsection{Convolution}
-The convolution utilize the \textit{convolve2d} function
+The convolution utilizes the \textit{convolve2d} function
 from Scipy to estimate the value of a current
@@
 The heuristic gathers all intermediate values into
-an histogram (using the \textit{histogram}
+a histogram (using the \textit{histogram}
 function of numpy) then give a weight to each
@@
 the board. Figure~\ref{fig:dot-representation}
-focus on the dot player and show the
+focuses on the dot player and show the
 transformation applied to the board for this
@@
 the rows, another for the columns and one for each
-diagonal. These convolution generate four smaller
+diagonal. These convolutions generate four smaller
 matrices showed in
 Figure~\ref{fig:convolution-results} (left side).
-The matrices are then flatened, concateneted and the
-absolute value is applied to every cell. Finally
+The matrices are then flattened, concatenated and the
+absolute value is applied to every cell. Finally,
 the histogram is built out of this matrix and each
@@
 
-This approach is strong for many reason. It is
+This approach is strong for many reasons. It is
 faster than calling python code because Scipy and
-Numpy rely on C writen functions. Uncomplete rows
-that contain zeros are more valued than uncomplete
+Numpy rely on C written functions. Incomplete rows
+that contain zeros are more valued than incomplete
 rows that contain opposite values. Finally,
 because the convolution mask is applied to empty
-spaces, the heuristic can build complexe
+spaces, the heuristic can build complex
 structures with space in the middle or above.
@@
 		\caption{The four matrices obtained after applying the four masks on the matrix that represent the dot, in Figure~\ref{fig:dot-representation}.
-						This Figure also represents the conversion to an histogram and how this histogram is used to find the heuristic value for the dot player.}
+						This Figure also represents the conversion to a histogram and how this histogram is used to find the heuristic value for the dot player.}
 		\label{fig:convolution-results}
@@
 symbols. Similarly to the previous heuristic,
-Vspace count the size of the lines, but for every
+\textit{Vspace} counts the size of the lines, but for every
 line connected to an empty cell, it uses the
@@
 the line is split in two with an empty cell in the
-middle, the combine value of both side gives the
+middle, the combined value of both side gives the
 value of the space. This is the reason why the two
@@
 the maximum value from them. The heuristic split the
-space in two categories: the space available in
-the next turn and the spaces too high to be reach
-with one card.
+space into two categories: the space available in
+the next turn and the spaces too high to be
+reached with one card.
 
-Finally, an histogram for each category is made and
+Finally, a histogram for each category is made and
 each value is multiplied by a weight then summed.
 Note that the two categories have different
-weights, because a space connected to a line of
-three symboles does not hold the same value if you
+weights because a space connected to a line of
+three symbols does not hold the same value if you
 can play on it on the next turn or not.
@@
 \end{figure}
+
+The default weights of \textit{Vspace} were hinted
+by a genetic algorithm~\cite{genetic}. The DEAP~\cite{deap}
+library was used to improve the initial weights
+which were a simple guess.  New weights would
+compete against a \textit{Vspace} heuristic with
+the original weights and the winning rate was used
+to evaluate them. In case the winning rate was
+zero, the new canditates were sorted by the
+average length of their game: the longer, the
+better. After a night of simulation, the results
+indicate that lines should not be counted except
+for lines of four symbols.  This change,
+greatly improved the winning rate of
+\textit{Vspace}.
+
 \clearpage
 \section{Difficulties}
-\begin{itemize}
-		\item Slowness of python, (numpy first)
-		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)
-\end{itemize}
+The biggest difficulty encountered during this
+project was the slowness of python. In the early
+stage of the project, the code was
+too inefficient and could not check more than one
+move in advance within the limit of three seconds.
+
+\textit{cProfile}~\cite{cprofile} is a module to profile
+python code. In particular,
+it gives the amount of time spent in each function
+called. This profiler was used to locate the most
+time-consuming function, then this information
+helps to restructure the code and enhance the
+performance.
+
+However, the results were hardly perceivable and
+the code remains slow even with the later use of
+optimized python libraries (Numpy and Scipy
+\cite{numpy}). Functions from these libraries rely
+on built-in functions which are written in C, then
+compiled into python module.  Therefore, the use
+of this type of function should improve the speed
+of the engine.
+
+However, none of the approaches above provide significant improvement. Finally,
+to reach a decent executing time for the tournament, one last step was taken
+and a python module written in C was developed.
+Figure~\ref{fig:performance-magic} shows the
+evolution of the performances as the most time-consuming function given by \textit{cProfile} were adapted to call the
+module. However, the python code remains and it is
+executed by default. 
+Figure~\ref{fig:performance-magic} shows that the
+MinMax algorithm is able to check four moves in
+advance after half of the key part used the
+module.
+
+\begin{figure}
+		\centering
+		\includegraphics[width=0.8\textwidth]{perfromances.png}
+		\caption{The amount of time (in seconds) taken
+		to run the MinMax algorithm on a board with the
+		card "5 B 1" already played. The MinMax
+		algorithm checked up to four moves in advance and
+		it used the Vspace heuristic described in
+		Section~\ref{sec:vspace}. The left column
+		shows the full python execution time (more
+		than 4 minutes) and the other columns show the
+		execution time as part of the code start using
+		the C module. The third column from the left,
+		"checking a move", indicate the time taken
+		when the heuristic function and the function
+		that check the moves use both the C module.}
+		\label{fig:performance-magic}
+\end{figure}
+
+\clearpage
 \section{Analyze}
-\begin{itemize}
-		\item How is the pruning depth 3 with a
-				recycling moves: no prun = 26s, prun=0.3s.
-		\item How are the performance at the tournament.
-\end{itemize}
+This section presents three analyses. A comparison
+between the two heuristics; a brief  overview of
+the alpha-beta pruning impact; and an experiment
+where the branching factor is limited in order to
+go deeper.
 
-\section{A Subsection Sample}
-Please note that the first paragraph of a section or subsection is
-not indented. The first paragraph that follows a table, figure,
-equation etc. does not need an indent, either.
+Before exploring these analyses, a brief
+explanation about winning rate tables. When two
+heuristics/algorithms are compared, two situations
+are considered: one algorithm
+play first in the first situation then the roles
+are inverted in the next situation. A quick experiment with the
+same algorithm and the same heuristic showed that
+the winning rate is higher for the second player.
+Therefore, we need to compare two situations.
 
-Subsequent paragraphs, however, are indented.
+\subsection{Heuristic comparison}
+In this section, we compare the efficiency between
+the two heuristics. Table~\ref{tab:win-heuristic}
+shows the winning rate of each heuristic out of 20
+games. \textit{Vspace} is used with a depth of
+four (four moves in advance) while the convolution
+only uses a depth of three. This difference in
+depth appears because the \textit{Vspace}
+heuristic is more optimized. Both heuristics are
+used by the tweaked \textit{MinMax} algorithm described
+above.
 
-\subsubsection{Sample Heading (Third Level)} Only two levels of
-headings should be numbered. Lower level headings remain unnumbered;
-they are formatted as run-in headings.
+The first row of Table~\ref{tab:win-heuristic}
+indicates that the \textit{Vspace} heuristic wins
+70 \%  of the game when it starts to play while the
+convolution heuristic wins the other 30 \%.
+From this table, we see that the Vspace heuristic
+performs better because it also wins all games
+when the convolution heuristic plays first.
 
-\paragraph{Sample Heading (Fourth Level)}
-The contribution should contain no more than four levels of
-headings. Table~\ref{tab1} gives a summary of all heading levels.
+To investigate the efficiency of the heuristic
+rather than the combination of its performance and
+efficiency, the previous experiment was redone
+with both heuristics using the same depth.
+Table~\ref{tab:win-heuristic-same} shows the
+winning ratio for this second experiment. We
+denote that the \textit{Vspace} heuristic is more
+efficient than the convolution.
 
 \begin{table}
-\caption{Table captions should be placed above the
-tables.}\label{tab1}
+	\begin{subtable}{0.5\textwidth}
+		\centering
+		\caption{Optimal depth}\label{tab:win-heuristic}
 		\begin{tabular}{|l|l|l|}
 		\hline
-Heading level &  Example & Font size and style\\
+		First player &  Vspace & Convolution\\
 		\hline
-Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
-1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
-2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
-3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
-4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
+		 Vspace 		 &  70 \% & 30 \%\\
+		 Convolution &  100 \% & 0 \%\\
+		\hline
+		\end{tabular}
+	\end{subtable}
+	\begin{subtable}{0.49\textwidth}
+		\centering
+		\caption{Same depth}\label{tab:win-heuristic-same}
+		\begin{tabular}{|l|l|l|}
+		\hline
+		First player &  Vspace & Convolution\\
+		\hline
+		 Vspace 		 &  55 \% & 45 \%\\
+		 Convolution &  100 \% & 0 \%\\
+		\hline
+		\end{tabular}
+	\end{subtable}
+	\caption{Shows the winning ratio between the two
+		heuristics. On the left, both heuristics use
+		the maximum depth they could reach. On the
+		right, they use the same depth (three moves
+		in advance). The column \textit{First player}
+		indicates which heuristic is playing first.}
+\end{table}
+
+\subsection{Alpha-Beta}
+The impact of Alpha-Beta pruning is important. A
+quick experiment with the \textit{Vspace}
+heuristic on depth 4 showed that the alpha-beta
+pruning speed-up the algorithm by a factor of 26
+(from 43 seconds to 1.6 seconds). The efficiency
+of the pruning is also seen in the number of time the
+heuristic function is called: 90000 calls with the
+pruning and more than 4 million without.
+Therefore, we can conclude that the Alpha-Beta
+pruning is a key element of the \textit{MinMax}.
+
+\subsection{Limiting the width}
+During the tournament, I heard that a team added a
+limitation in width in addition to the depth
+limitation.
+Therefore, I decided to evaluate this idea in my
+own project. I design a new version of the MinMax
+algorithm where at the beginning of each function
+(\textit{min} and \textit{max}) each move is rated
+with the heuristic and only the 25 best are
+explored. This modification fixed the branching
+factor at 25 and allows the algorithm to go
+one level deeper. The factor of 25 was choosen
+because it allows going just one level deeper
+within the time limit while still exploring many
+moves.
+
+Table~\ref{tab:win-limit-width} shows the win rate
+between the \textit{MinMax} with unlimited width
+and the one with a branching factor of 25. The Table
+reveals that the limited search with one level
+deeper has a poor winning rate whether it starts
+to play or not.
+Therefore, in this project with the
+\textit{Vspace} heuristic, limiting the width is inefficient
+and does not give an advantage.
+\begin{table}
+		\centering
+		\caption{Win ratio when the width is limited.}\label{tab:win-limit-width}
+		\begin{tabular}{|l|l|l|}
+		\hline
+		First player &  Limited & Unlimited\\
+		\hline
+		 Limited &  0 \% & 100 \%\\
+		 Unlimited &  35 \% & 65 \%\\
 		\hline
@@
 
-\noindent Displayed equations are centered and set on a separate
-line.
-\begin{equation}
-x + y = z
-\end{equation}
-Please try to avoid rasterized images for line-art diagrams and
-schemas. Whenever possible, use vector graphics instead (see
-Fig.~\ref{fig1}).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig1.eps}
-\caption{A figure caption is always placed below the illustration.
-Please note that short captions are centered, while long ones are
-justified by the macro package automatically.} \label{fig1}
-\end{figure}
-
-\begin{theorem}
-This is a sample theorem. The run-in heading is set in bold, while
-the following text appears in italics. Definitions, lemmas,
-propositions, and corollaries are styled the same way.
-\end{theorem}
-%
-% the environments 'definition', 'lemma', 'proposition', 'corollary',
-% 'remark', and 'example' are defined in the LLNCS documentclass as well.
-%
-\begin{proof}
-Proofs, examples, and remarks have the initial word in italics,
-while the following text appears in normal font.
-\end{proof}
-For citations of references, we prefer the use of square brackets
-and consecutive numbers. Citations using labels or the author/year
-convention are also acceptable. The following bibliography provides
-a sample reference list with entries for journal
-articles~\cite{ref1}, an LNCS chapter~\cite{ref2}, a
-book~\cite{ref1}, proceedings without editors~\cite{ref1},
-and a homepage~\cite{ref_2}. Multiple citations are grouped
-\cite{ref1,ref2}.
-% ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
+\section{Conclusion}
+This project taught me many things about AI
+programming for two player game. It showed me how
+the weight of a heuristic can be optimized. It
+proved me that going deeper in the state tree
+is not the better option if all possibilities are
+not properly explored. Finally, I learned to program python
+module and to combine the performance of C with
+the simplicity of python when performances are
+required.
 \bibliographystyle{splncs04}

### 2019-03-28 src/min_max_player.py
Last minute commit

@@ -26,7 +26,9 @@ class MinMaxPlayer:
         if isinstance(sort_moves, bool):
             self.sort_moves = sort_moves
         else:
-            self.sort_moves = (sort_moves == 1)
+            self.sort_moves = (int(sort_moves) == 1)
+        self.lili = 25
+        print(self.name, self.sort_moves)
 
     @timeit
     def play(self, id_ai, engine):
@@ -88,7 +90,7 @@ class MinMaxPlayer:
         s = sum([mv.recycling for mv in moves])
         if self.sort_moves:
             value = [self.value_move(engine, mv, self.id_ai) for mv in moves]
-            moves = [mv[1] for mv in sorted(value, key=itemgetter(0), reverse=True) if mv[0] > -900000 ]
+            moves = [mv[1] for mv in sorted(value, key=itemgetter(0), reverse=True) if mv[0] > -900000 ][:self.lili]
 
         best_val = -100000
         best_move = None
@@ -132,7 +134,7 @@ class MinMaxPlayer:
         if self.sort_moves:
             value = [self.value_move(engine, mv, self.id_ai) for mv in moves]
             best_opponent = min([mv[0] for mv in value])
-            moves = [mv[1] for mv in sorted(value, key=itemgetter(0), reverse=True)]
+            moves = [mv[1] for mv in sorted(value, key=itemgetter(0), reverse=True)][:self.lili]
 
         #Does the TA or the teacher of AI read this comment?
 

### 2019-03-28 src/practice.py

@@ -7,7 +7,7 @@ from random_player import RandomPlayer
 from min_max_player import MinMaxPlayer
 from demo_minmax import DemoMinMaxPlayer
 from monte_carlo import MonteCarloPlayer
-from heuristics import HeuristicConvolution, HeuristicNeuralNetwork, HeuristicBasic, HeuristicVspace, HeuristicRandom, HeuristicDemo
+from heuristics import HeuristicConvolution, HeuristicNeuralNetwork, HeuristicBasic, HeuristicVspace, HeuristicRandom, HeuristicDemo, HeuristicMuggleVspace
 from sklearn.metrics import accuracy_score
 import cProfile
 import pstats
@@ -135,39 +135,38 @@ p = MinMaxPlayer(h, 4, False)
 p2 = MinMaxPlayer(h)
 engine.initialize_player(p2, p)
 play_move(engine, "0 5 B 1")
-play_move(engine, "0 4 A 1")
-play_move(engine, "0 5 B 2")
-play_move(engine, "0 2 C 3")
-play_move(engine, "0 2 B 3")
-play_move(engine, "0 2 A 3")
-play_move(engine, "0 5 B 5")
-play_move(engine, "0 3 B 6")
-play_move(engine, "0 1 B 7")
-play_move(engine, "0 8 A 5")
-play_move(engine, "0 5 F 1")
-play_move(engine, "0 8 A 7")
-play_move(engine, "0 4 G 2")
-play_move(engine, "0 4 C 8")
-play_move(engine, "0 6 A 9")
-play_move(engine, "0 2 B 8")
-play_move(engine, "0 6 A 11")
-play_move(engine, "0 6 C 10")
-play_move(engine, "0 2 B 10")
-play_move(engine, "0 1 B 12")
-play_move(engine, "0 4 G 4")
-play_move(engine, "0 8 G 6")
-play_move(engine, "0 6 F 2")
-play_move(engine, "0 4 F 4")
-play_move(engine, "A 11 A 12 6 E 1")
-play_move(engine, "A 9 A 10 8 E 3")
-play_move(engine, "A 7 A 8 6 A 7")
-play_move(engine, "B 12 C 12 8 A 9")
-play_move(engine, "E 3 E 4 2 E 3")
-play_move(engine, "B 10 B 11 8 E 5")
-# play_move(engine, "0 6 F 1")
+# play_move(engine, "0 4 A 1")
[…]

### 2019-03-28 src/main.py
Add muggle vspace in main.p in main.py

@@ -7,7 +7,7 @@ from random_player import RandomPlayer
 from min_max_player import MinMaxPlayer
 from demo_minmax import DemoMinMaxPlayer
 from monte_carlo import MonteCarloPlayer
-from heuristics import HeuristicConvolution, HeuristicNeuralNetwork, HeuristicBasic, HeuristicVspace, HeuristicRandom, HeuristicDemo
+from heuristics import HeuristicConvolution, HeuristicNeuralNetwork, HeuristicBasic, HeuristicVspace, HeuristicRandom, HeuristicDemo, HeuristicMuggleVspace
 from sklearn.metrics import accuracy_score
 import cProfile
 import pstats
@@ -65,8 +65,7 @@ def load_random_player(l):
     return RandomPlayer(*l)
 def load_vspace_heuristic(dm):
     if not dm:
-        print("Using vspace heuristic require dark magic.")
-        raise NameError("No dark magic")
+        return HeuristicMuggleVspace()
     return HeuristicVspace()
 
 #define the run functions

### 2019-03-28 src/min_max_player.py

@@ -27,8 +27,7 @@ class MinMaxPlayer:
             self.sort_moves = sort_moves
         else:
             self.sort_moves = (int(sort_moves) == 1)
-        self.lili = 25
-        print(self.name, self.sort_moves)
+        self.lili = 25 #Width limitation when sorting.
 
     @timeit
     def play(self, id_ai, engine):
